---
title: "MATH1324 Assignment 1"
subtitle: Modeling Body Measurements
output:
html_notebook: default
---

## Student Details

Simon Karumbi (s3455453)

## Problem Statement

The goal of this investigation is to determine whether one of the body measurements from the sampled data of physically active males and females, fits a normal distribution. This will help determine the best strategy to purchase a sizes for an activewear brand also using a normal distribution to determine which waist sizes should be bought into.

As we wish to test a model for clothing, waist girth was selected as a variable where a normal distribution would likely fit for both men and women, while also being an easy to implement purchasing strategy for sizing of bottom wear.

## Load Packages

```{r}
library(forcats)
library(dplyr)
library(ggplot2)
library(readr)
```

## Data

Import the body measurements data and prepare it for analysis. Show your code.


```{r}
# Reading in flat file
bdims <- read_csv("bdims.csv")

```


## Summary Statistics

Calculate descriptive statistics (i.e., mean, median, standard deviation, first and third quartile, interquartile range, minimum and maximum values) of the selected measurement grouped by sex.

```{r}
# Factoring and labelling 'Sex' variable to better understand the information
bdims$sex <- bdims$sex %>% as.factor() %>% fct_recode("Female"="0", "Male"="1")

# Creating a manageable dataframe for the chosen variable (Waist girth measured at the narrowest part of torso below the rib cage as average of contracted and relaxed position)
df <- data.frame(Age = bdims$age, Sex = bdims$sex, Height = bdims$hgt, Weight = bdims$wgt, Waist = bdims$wai.gi)

# Summarising the mean, median, min, max, standard deviation, first and third quartile, count and missing count of the waistngirth data for males and females
df %>% group_by(Sex) %>% summarise(
Min = min(Waist,na.rm = TRUE),
Max = max(Waist,na.rm = TRUE),
Median = median(Waist, na.rm = TRUE),
Mean = mean(Waist, na.rm = TRUE),
SD = sd(Waist, na.rm = TRUE), 
Q1 = quantile(Waist,probs = .25,na.rm = TRUE),
Q3 = quantile(Waist,probs = .75,na.rm = TRUE),
Count = n(), 
Missing = sum(is.na(Waist)))

```


## Distribution Fitting

Compare the empirical distribution of selected body measurement to a normal distribution separately in men and in women. You need to do this visually by plotting the histogram with normal distribution overlay. Show your code.

```{r}
# Data frames to group Male and Female data for easier manipulation and visualisation
df_male <- df %>% filter(Sex == "Male")
df_female <- df %>% filter(Sex == "Female")

# Assigning variables for Men's and Women's mean and sd for calculating the normal distribution
male_mean <- mean(df_male$Waist)
male_sd <- sd(df_male$Waist)
female_mean <- mean(df_female$Waist)
female_sd <- sd(df_female$Waist)

# Histogram of Men's waist girth in centimeters with a normal distribution overlay
hist(df_male$Waist, probability = TRUE, xlab = "Waist girth (cm)", main = "Waist measurement in Males", col = "lightblue", xlim = c(65,115), breaks = 20)
x <- c(65:115)
y <- dnorm(x = x, mean = male_mean, sd = male_sd)
lines(x = x, y = y, col = "maroon", lwd=2)

# Histogram of Women's waist girth in centimeters with a normal distribution overlay
hist(df_female$Waist, probability = TRUE, xlab = "Waist girth (cm)", main = "Waist measurement in Females", col = "pink", xlim = c(55,105), breaks = 20)
x <- c(55:105)
y <- dnorm(x = x, mean = female_mean, sd = female_sd)
lines(x = x, y = y, col = "maroon", lwd =2)
```

## Interpretation

Male waist measurement fits a normal distribution well with a slight positive skew. Female waist measurement also moderately fits a normal distribution with a more positive skew. It can be said that both male and female waist girth measurements approximately fit a normal distribution. 

The fact that both male and female waist girths approximately fit a normal distribution makes understanding size requirements for an activewear brand fairly easy to predict using a normal distribution model. Bottom wear size runs purchased using a normal distribution will most likely benefit the brand financially through increased sales through increased availability.

